---
title: "Exercise Solutions Module 4"
author: "Alex Thomson - Stats4SD"
date: "30/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(learnr)
Happiness <- read.csv("Happiness.csv")
```

## Exercises
**Exercise 1: Now that you have imported the Happiness dataset into R – let’s do a quick exercise to recap what we learnt in Modules 1-3, but writing out the code entirely in RStudio instead of online.**

1.	Find the 5 countries with the lowest ‘generosity scores’  

We of course need to start with `arrange()` so we can order our data according to the Generosity variable. We do not need to add in the `desc()` function because we want the lowest scores to come up first.

We don't have to but we can then use `select()` to select only the Country.or.region and Generosity columns as for this example, these are the only variables we are focusing on.

Finally we can filter down to just those 5 lowest scores by using the `slice()` function. This can work in a similar way to `select()`function except we are selecting rows rather than columns. We can use a list of row numbers in `c()` or if the rows are grouped together we can use the colon in the same way we use it in the `select()` function. Here we use `1:5` to singal we want to only keep rows 1 through 5.

```{r ex1}
Happiness%>%
  arrange(Generosity)%>%
  select(Country.or.region, Generosity)%>%
  slice(1:5)
```

2.	Produce a subset of countries which have below average scores for both GDP and generosity.  

Here we can just use `filter()` in similar ways that we have done before such as when looking for a lowest or highest rated movie as in the previous module. But rather than using `min` or `max` we can use `mean` instead to filter down to those that are below the average for a given variable. We do this for both Generosity and GDP, remember to use `&`. We have also specified `na.rm = TRUE` to correct for any missing data, there shouldn't be in this data but it is often a good idea to include this argument to be on the safe side.

```{r ex2}
Happiness_subset <- Happiness%>%
  filter(
    Generosity < mean(Generosity, na.rm = TRUE) & GDP.per.capita < mean(GDP.per.capita, na.rm = TRUE)
  )

Happiness_subset
```


3.	Make a plot showing the relationship between generosity and GDP per capita  

We start by setting up our aesthetics in the usual way. It does not really matter either way which variable is on which axis for this case.

As we are investigating the relationship between two continuous variables, our preferred choice of plot would be a scatterplot. For which the funciton is `geom_point()`. We do not need to specify any further arguments.

I can't really tell anything directly, it doesn't seem there is much of a relationship just by looking at the points. Alternatively i can add in a line of best fit using `geom_smooth()`. Now here we have specified the argument method to be equal to "lm". The method argument tells R how to plot this line of best fit. By default smooth tries to pick itself what is the best method, usually opting for what is known as "loess" which is a sort of moving average. "lm" however will tell R to fit the line using a more standard "linear model", or in other words regression. 

There is a slight downwards trend to the line but this may not be substantial enough to conclude there is a relationship.
 
```{r}
Happiness%>%
  ggplot(aes(x = Generosity, y = GDP.per.capita))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "GDP per capita by National Generosity Score")
```
 

**Exercise 2: Download the file below that contains the imdb dataset and the solutions for the module 3 exercises. Save these in your project folder. Import the data using RStudio menu and go through the R Markdown file to reproduce the solutions in your own machine**

[Module-3-Data-and-Solutions.zip](https://github.com/stats4sd/r2020_04Quiz/raw/main/Module-3-Data-and-Solutions.zip)

These exercises were largely reviewed in the in-person session as they are a review of Module 3 exercises but using Rstudio rather than the online workbooks.